---
title: "Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all()
script_path <- system.file("vignette_helpers", "db_functions.R", package = "curveRmetrics")
if (file.exists(script_path)) {
  source(script_path)
} else {
  warning("Script file not found at expected path: ", script_path)
}
conn <- get_db_connection()

freq_curves_df  <- get_freq_curves(conn, "curveTest2")
bayes_curves_df <- get_bayes_curves(conn, "curveTest2")
bayes_param_ci_df <- get_bayes_param_ci(conn, "curveTest2")
freq_param_ci_df <- get_freq_parameters(conn, "curveTest2")
standards <- get_standards(conn, "curveTest2")

```


```{r}
# combine both the Frequentest and Bayesian curves together for quality metrics
# calculations and comparisons. 
curves_df <- rbind(freq_curves_df, bayes_curves_df)
# naturalize the parameters
curves_nat <- curves_df %>%
  transform_to_natural_units()

head(curves_nat)
```

Select a specific `curve_id` to visualize comparisions. The `curveRmetrics`
package will compute Quality Control (QC) metrics for each `curve_id` in the 
naturalized parameter set (`curves_nat`).

## Compute Inflection Point 

```{r}
nat_params <- compute_inflection_point(curves_nat, verbose = F)
head(nat_params)
```

## Assay Sensitivity 
```{r}
nat_sens <- compute_assay_sensitivity(nat_params, verbose = T)
head(nat_sens)
```

## Limits of Detection

First the raw confidence/credible intervals and parameters are transformed to natural units and combined together in one dataset
for comparison.
```{r}
bayes_param_ci_nat <- transform_ci_to_natural_units(bayes_param_ci_df)
freq_param_ci_nat <- transform_ci_to_natural_units(freq_param_ci_df)

params_ci_nat <- rbind(freq_param_ci_nat, bayes_param_ci_nat)
params_ci_nat
```

```{r}
lods_nat <- generate_lods(param_ci_df = params_ci_nat, verbose = T)
lods_nat
```

## Reliable Detection Limits

```{r}
rdl_nat <- compute_mdc_rdl(param_ci_df = params_ci_nat, lods = lods_nat, verbose = F)
head(rdl_nat)
```

## Limits of Quantification
### Curvature-Based

```{r}
d2_natural <- compute_second_deriv_df(curves_df = curves_nat)
head(d2_natural)
```

```{r, fig.width = 14, fig.height = 8, out.width = "100%"}
curve_id <- 74238

compare_second_derivative(second_derivative_df = d2_natural, curve_id = curve_id)
```

Calculate the shaped-based Limits of Quantification (LOQs) from the second derivative

```{r}
loqs_nat <- compute_loqs(curves_nat, second_deriv_df = d2_natural, verbose = F)
head(loqs_nat)
```

## Attach Quality metrics to the original curve dataframe for plotting and uploading back to database.
```{r}
curves_quality_df <-attach_quality_metrics(curves_df = curves_df,
                                           lods = lods_nat, 
                                           rdls = rdl_nat,
                                           sensitivity = nat_sens,
                                           loqs = loqs_nat, 
                                           inflection = nat_params)

head(curves_quality_df)
```


```{r, fig.width = 14, fig.height = 10, out.width = "100%"}
compare_quality_metrics(curves_nat, standards, curves_quality_df, curve_id = curve_id)
```




