---
title: "Computing Standard Curve Quality Control Metrics using curveRmetrics" 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computing Standard Curve Quality Control Metrics using curveRmetrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette demonstrates the end-to-end `curveRmetrics` workflow for evaluating
standarded curve quality. Starting from raw Frequentist and Bayesian curve fits
stored in a database, we derive a standardized set of Quality Control (QC) metrics:

- **Inflection point** — The point on the standard curve where the concavity transitions from concave up to concave down. It is the point where the assay is most sensitive to measurement errors in the measured response of the assay.

- **Assay sensitivity** — Measured by the slope of the inflection point.

- **Limits of Detection (LODs)** — Lower and upper LODs are defined as the upper 97.5% confidence bound of the lower asymptote and the lower 2.5% confidence bound of the upper asymptote, respectively (Rajam et al.). 

- **Reliable Detection Limits (RDLs)** - Lower RDL: The lowest concentration at which the assay consistently produces a signal above background with 95% confidence based on the fit of the standard curve (Rajam et al.). Upper RDL: Analogously, the highest concentration at which the assay consistently produces a signal below the upper asymptote (saturation) with 95% confidence, based on the fit of the standard curve.

- **Minimum Detectable Concentration** - The smallest antibody concentration that produces a signal the assay can detect above background (Rajam et al.). This corresponds to the x-coordinate of the Lower Limit of Detection in the legend, as it is on the concentration axis.

- **Limits of Quantification (LOQ)** - Defines a region of assay response (MFI) and concentration where sample estimates have less measurement error. Limits of Quantification are derived from the local minimum and maximum of the second derivative of x given y of the standard curve (Daly et al.), (Jeanne L Sebaugh and P. D. McCray) , (Sanz et al.).


All metrics are computed in natural (back-transformed) parameter units so that
results are directly interpretable and comparable across fitting methods.


## Setup: Load Curve Data from Database

We load Frequentist and Bayesian curve fits for the `curveTest2` study containing data from ELISA immunoassays,
along with their parameter confidence/credible intervals and assay standards. 
Other studies or similar fit data from standard curves can be used as input as long as they follow the functional forms from accompanying `curveRfreq` or `stanassay` R packages. Curve fits from multiplex bead arrays and other immunoassays with a standard curve can be used. 

```{r setup}
devtools::load_all()
script_path <- system.file("vignette_helpers", "db_functions.R", package = "curveRmetrics")
if (file.exists(script_path)) {
  source(script_path)
} else {
  warning("Script file not found at expected path: ", script_path)
}
conn <- get_db_connection()

freq_curves_df  <- get_freq_curves(conn, "curveTest2")
bayes_curves_df <- get_bayes_curves(conn, "curveTest2")
bayes_param_ci_df <- get_bayes_param_ci(conn, "curveTest2")
freq_param_ci_df <- get_freq_parameters(conn, "curveTest2")
standards <- get_standards(conn, "curveTest2")

```


Curve fit results from both Frequentist and Bayesian fitting methods are combined into a single data frame so that all downstream QC functions operate on Frequentist and Bayesian curves simultaneously, enabling direct method comparisons.

```{r}
# combine both the Frequentest and Bayesian curves together for quality metrics
# calculations and comparisons. 
curves_df <- rbind(freq_curves_df, bayes_curves_df)
# naturalize the parameters
curves_nat <- curves_df %>%
  transform_to_natural_units()

head(curves_nat)
```

**Note:** `transform_to_natural_units()` back-transforms four curve parameters (`a`, `b`, `c`, `d`) from their fitted scales to natural units.

Select a specific `curve_id` to visualize comparisons. The `curveRmetrics`
package will compute Quality Control (QC) metrics for each `curve_id` in the 
naturalized parameter set (`curves_nat`).

The `verbose` argument in the `curveRmetrics` shows QC output messages. 
To hide these messages set `verbose` to `FALSE`.

## Compute Inflection Point 

```{r}
nat_params <- compute_inflection_point(curves_nat, verbose = F)
head(nat_params)
```

## Assay Sensitivity 
```{r}
nat_sens <- compute_assay_sensitivity(nat_params, verbose = T)
head(nat_sens)
```

## Limits of Detection

First the raw confidence/credible intervals and parameters are transformed to natural units and combined together in one dataset
for comparison.
```{r}
bayes_param_ci_nat <- transform_ci_to_natural_units(bayes_param_ci_df)
freq_param_ci_nat <- transform_ci_to_natural_units(freq_param_ci_df)

params_ci_nat <- rbind(freq_param_ci_nat, bayes_param_ci_nat)
params_ci_nat
```

The limits of detection are then calculated.
```{r}
lods_nat <- generate_lods(param_ci_df = params_ci_nat, verbose = T)
lods_nat
```

## Reliable Detection Limits

```{r}
rdl_nat <- compute_mdc_rdl(param_ci_df = params_ci_nat, lods = lods_nat, verbose = F)
head(rdl_nat)
```

## Limits of Quantification
### Curvature-Based Approach

We first compute the second derivative across the concentration range for all curves then select a `curve_id` of interest for comparision.:
```{r}
d2_natural <- compute_second_deriv_df(curves_df = curves_nat)
head(d2_natural)
```

```{r, fig.width = 14, fig.height = 8, out.width = "100%"}
curve_id <- 74238

compare_second_derivative(second_derivative_df = d2_natural, curve_id = curve_id)
```

We then calculate the shaped-based Limits of Quantification (LOQs) from the second derivative profile for all curves:

```{r}
loqs_nat <- compute_loqs(curves_nat, second_deriv_df = d2_natural, verbose = F)
head(loqs_nat)
```

## Quality Control Summary amd Comparision


All quality control metrics computed above are attached to the original curve data frame. The enriched object (`curves_quality_df`) is suitable for both visualization
and uploading results back to the database.

```{r}
curves_quality_df <-attach_quality_metrics(curves_df = curves_df,
                                           lods = lods_nat, 
                                           rdls = rdl_nat,
                                           sensitivity = nat_sens,
                                           loqs = loqs_nat, 
                                           inflection = nat_params)

head(curves_quality_df)
```

Quality Control thresholds and limits for a curve can be plotted enabling a comparison of Frequentist and Bayesian fits. 

```{r, fig.width = 14, fig.height = 10, out.width = "100%"}
compare_quality_metrics(curves_nat, standards, curves_quality_df, curve_id = curve_id)
```




